Energy and exergy analysis of a cruise ship

### 2018-07-04 Manuscript/Manuscript.tex
Addition of the abstract and of the nomenclature

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
%% The amsthm package provides extended theorem environments
\usepackage{amsthm}
%% Tthe amsmath package provides the "align" environment for nicer equation formatting
\usepackage{amsmath}
%% The package tabls helps in having nicer tables with higher spaces
\usepackage{tabls}
%% The package multirow allows merging together multiple cells vertifcally
\usepackage{multirow}
%% The package subcaption is for helping when using subfigures and their captions
\usepackage{subcaption}
%% This package is used for having nicer tables
\usepackage{booktabs}
% Package for improving table handling
\usepackage{array}
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
% Package for drawing pictures
\usepackage{tikz}
\usepackage{pgfplots}

@@

\author[EPFL]{Francesco Baldi}
\author[Linnaeus]{Fredrik Ahlgren}
\author[DTU]{Tuong-Van Nguyen}
\author[LTU]{Marcus Thern}
\author[CTU]{Karin Andersson}

\address[EPFL]{Industrial Process Energy Systems Engineering (IPESE), \'{E}cole Polytechnique F\'{e}d\'{e}rale de Lausanne, 1950, Sion, Switzerland}
\address[Linnaeus]{Kalmar Maritime Academy, Linnaeus University, Kalmar, Sweden}
\address[DTU]{Department of Mechanical Engineering, Technical University of Denmark, Lyngby, Denmark}
\address[LTU]{Energy Sciences, Lund University, Lund SE-22100, Sweden}
\address[CTU]{Department of Mechanics and Maritime Sciences, Chalmers University of technology, Gothenburg, Sweden}

\begin{abstract}
%% Text of abstract

@@

@@

CO$_2$ emissions from shipping in 2012 amounted to a total of 949 million tonnes, contributing to 2.7\% of global anthropogenic CO$_2$ emissions \cite{Smith2014}. Although this contribution appears relatively low, the trend is that shipping will play an even greater role in the future due to the increased transport demand according to all IMO future scenarios \cite{Smith2014}. As an example, global transport demand has increased by 3.4\% in 2014, compared to a global GBP growth of 2.5\% the same year, which shows how shipping tends to rise even faster than global economy \cite{UNCTAD2015}. The OECD countries have reduced the CO$_2$ impact from shipping, but a larger amount has been moved to the non-OECD countries [3]. The fact that shipping needs to even further reduce its CO$_2$ emissions in the near future is essential for being able to achieve the goals of maintaining the climate below a 2-degree level in 2050 [4].

In addition to considerations related to GHG emissions, an emission control area (ECA) is enforced in the Baltic Sea  by the International Maritime Organisation since January 2015. This ECA stipulates that the fuel used must not contain more than 0.1\% sulphur, therefore requiring the use of more expensive distillate fuels. More generally making shipping sustainable is a challenge that will demand growing attention by the shipping industry \cite{Andersson2016}

@@

Altogether, these conditions present a challenge to the shipping companies that attempt to reduce their fuel consumption, environmental impact, and operative costs. A wide range of fuel saving solutions for shipping are available and partially implemented in the existing fleet, both from the design and operational perspectives; several specific studies have been conducted on these technologies, and a more detailed treatise would be out of the scope of this work. In this context, it has been acknowledged that the world fleet is heterogeneous, and measures need to be evaluated on a ship-to-ship basis \cite{Bouman2017}. In this process, a deeper understanding of energy use on board of the specific ship is vital.

@@

Most authors focused on the propulsion part of the problem, as this is often the most relevant energy demand on board. Shi et al. \cite{Shi2009} proposed a modeling approach for predicting ship fuel consumption of a cargo/passenger ferry; Theotokoats and Tzelepis \cite{Theotokatos2015} applied a similar procedure to the case of a Handymax product carrier, similarly to Tillig et al. \cite{Tillig2016}, who also added the dynamic element to their predictive model.

@@

% Data-driven analysis
Other authors filled the gap by both including electric energy demand, and by basing their work on measurements from actual ship operations. The work presented by Thomas et al. \cite{Thomas2010} and Basurko et al. \cite{Basurko2013} shows the application of energy auditing methods to fishing vessels. This approach represents a step forward towards improving the understanding of how energy is used on board during actual ship operations.

@@

% Heat availability
Heat demand is rarely a subject of concern on board ships, with some notable exceptions. This is due to a combination of generally low demand and high availability from the waste heat of the engines. In a previous publications by the authors \cite{Baldi2015a}, for instance, it is shown that in the case of a product tanker, although the heat demand was estimated to account for roughly 20\% of the total energy demand of the ship on a yearly basis, it only contributes to 4.1\% of the total fuel consumption (contribution of the auxiliary boilers), while the rest of the demand is fulfilled using waste heat.

%% WHR and heat availability
It should be noted that, however, much research effort has been devoted, especially in recent times, to the improvement of the efficiency of ship energy systems by recovering the waste heat available from the engines. With reference to different types of technologies, case studies, and designs,the several authors showed the existance of a quite significant potential for energy saving when WHR systems are employed, ranging from around 1\% for single-pressure steam cycles applied to two-stroke engines \cite{Theotokatos2012} to more complex systems based on ORCs (up to 10\%, \cite{Hountalas2012}) or including the cooling systems as a source of waste heat (over 10\%, \cite{Dimopoulos2012}).  The case of the installation of an ORC on board of the vessel investigated in this study (see Section \ref{sec:met:case}) was presented in \cite{Mondejar2017}, showing a similar potential.

The potential uses for waste heat on board are not limited to improving the efficiency of the power plant. Waste heat is commonly used for fulfilling on board heat demand for spatial heating and freshwater generation \cite{Baldi2015a,Molland2011,Mccarthy1990};  Balaji et al. \cite{Balaji2012} proposed the use of waste heat for ballast-water systems ; Salmi et al. \cite{Salmi2017} suggested its use for adsorption refrigeration systems. A detailed review of potential uses for waste heat from marine engines is presented by Shu et al \cite{Shu2013}.

@@

% Exergy analysis
Exergy analysis provides a more accurate estimation of the potential for energy recovery on board. The application of exergy analysis to the case of ship energy systems is, however, still limited. Dimopoulos et al. \cite{Dimopoulos2012} showed how the process optimizing the WHR system of a container ship can be more efficient if exergy efficiency, rather than energy efficiency, is used as the target of the optimization; Baldi et al. \cite{Baldi2015a} also analyzed the exergy flows on board of a product tanker, showing that this allows having a more accurate understanding of what parts of the system show potential for improvement ; similar results were obtained by Marty et al. \cite{Marty2016}, who focused on the power plant of a cruise ship. Koroglu et al. \cite{Koroglu2017} made a step futher also including advanced exergy analysis in their study.

@@

In this paper, we present the application of energy and exergy analysis (described in Section \ref{sec:met:energyExergy}) to a cruise ship. This is done for one specific case study vessel, that is described in detail in Section \ref{sec:met:case}; in Section \ref{sec:met:gathering} we present the specifics of the available information, in particular the measured data and the technical documentation; the specific assumptions and methods used for processing the available information into  energy and exergy flows are specified in Section \ref{sec:met:processing}. In these regards, the estimation of the heat demand is treated in a separate section (Section \ref{sec:met:heat}) as it constitutes a particularly challenging task and is one of the central aspects of this article.

@@

\begin{align}
	\sum_{\mathrm{in}} \dot{H}_{\mathrm{in}} &= \sum_{\mathrm{out}} \dot{H}_{\mathrm{out}} \\
	\dot{H}_{\mathrm{fuel}} + \dot{H}_{\mathrm{air}} &= \sum_{{\mathrm{waste}}}\dot{H}_{{\mathrm{waste}}}+\dot{W}_{\mathrm{el}}+\dot{Q}_{\mathrm{heating}}
\end{align}

The left-hand side term represents, on a time rate basis, the energy associated with the fuel consumed in the boilers and engines ($\dot{H}_{\mathrm{fuel}}$) and the air used in the combustion processes ($\dot{H}_{\mathrm{air}}$). The right-hand side term denotes the power ($\dot{W}_{\mathrm{el}}$) and heat ($\dot{Q}_{\mathrm{heating}}$) required on-site (e.g. propulsion and fuel heating), and the heat discharged into the environment ($\sum_{{\mathrm{waste}}}\dot{H}_{{\mathrm{waste}}}$) with, for instance, the exhaust gases.

The energy flow associated with a material stream is calculated as the sum of the physical and chemical enthalpies, and kinetic and potential energies are neglected. The physical energy is taken as the relative enthalpy, as underlined in \cite{Kotas2013}, and the chemical energy is taken as the lower/higher heating value. The environmental conditions taken for the present analysis are the ambient pressure (1.01 bar) and seawater temperature (measured) .

@@

\begin{align}
	\sum_{\mathrm{in}} \dot{E}_{\mathrm{in}} &= \sum_{\mathrm{out}} \dot{E}_{\mathrm{out}} +\dot{E}_{d} \\
	\dot{E}_{\mathrm{fuel}} + \dot{E}_{\mathrm{air}} &= \sum_{{\mathrm{waste}}}\dot{E}_{{\mathrm{waste}}}+\dot{E}_{W}+\dot{E}_{Q,\mathrm{heating}}+\dot{E}_{d} 
\end{align}

The left-hand side term represents, on a time rate basis, the exergy associated with the fuel and air. The right-hand side term denotes the exergy of the waste streams, the exergy transfers with heat and power, and the exergy destroyed in the ship system. Kinetic and potential exergies are neglected, and the exergy of a given material stream is derived as the sum of the physical and chemical exergies. The chemical exergy is calculated based on the reference environment of \cite{Szargut1989}, and its value is approximatively equal to the higher heating value for hydrocarbon fuels. The exergy destruction can be calculated from the Gouy-Stodola theorem \cite{Kotas2013}.

The exergy balance may alternatively be formulated as:
\begin{equation}
\dot{E}_{\mathrm{p}} = \dot{E}_{\mathrm{f}} - \dot{E}_{\mathrm{d}} - \dot{E}_{\mathrm{l}} 
\end{equation}

where $\dot{E}_{\mathrm{p}}$ is called the exergy product, and corresponds to the desired output of the system, in exergy terms (for example, the power produced in an engine). $\dot{E}_{\mathrm{f}}$ denotes the exergy fuel, and represents the resources spent to drive the studied process (for instance, the fuel used in a combustion process). The last term $\dot{E}_{\mathrm{l}}$ corresponds to the losses of a system, such as the heat discharged into the environment with cooling water.

@@

@@

The ship systems are summarized in Figure \ref{fig:shipSystems}. The propulsion system is composed of two equal propulsion lines, each made of two engines, a gearbox, and a propeller. The main engines are four W\"{a}rtsil\"{a} 4-stroke Diesel engines (ME) rated 5850 kW each. All engines are equipped with selective catalytic reactors (SCR) for NO$_X$ emissions abatement. Propulsion power is needed whenever the ship is sailing; however, it should be noted that the ship rarely sails at full speed, and most of the time it only needs one or two engines operated simultaneously.

@@

Auxiliary power is provided by four auxiliary engines (AE) rated 2760 kW each. Auxiliary power is needed on board for a number of alternative functions, from pumps in the engine room to lights, restaurants, ventilation and entertainment for the passengers.

@@

The operational data was collected on board from the ships' machine logging and surveillance system. The data sampling rate was selected to 15 minutes, automatically calculated as averages over the respective periods by the on board data logging systems, as an acceptable compromise between accuracy and computational time required for the data processing. Data consistency was checked individually by creating a descriptive statistic and a histogram of each measurement. A filter was applied by setting up a pre-defined maximum and minimum value.

@@

The main principles applied to the modeling are presented in the following sections and summarized in Figure \ref{fig:dataprocessing}, while a more thorough description and validation of the different approaches is presented in the Appendix. The data is first processed to the main and auxiliary engines modules to derive the related energy flows. These are then used to calculate the electric power demand, the energy flows in the cooling systems and, consequently, the heat recovered in the HTHR systems. This information combined is used to determine the heating demand. Finally, all flows thus obtained are aggregated to determine energy and exergy flows and related efficiencies.

@@

The Diesel engines are the only power source of the system, subdivided in four main engines uniquely devoted to propulsion and four auxiliary engines used for on board electric power generation. The energy and exergy analysis of the engines is not limited to the boundary of the system, but is brought to the detail of the different components that constitute the engine systems, as shown in Figure \ref{fig:DieselEngines}.

@@

@@

For both auxiliary and main engines, part of the relevant internal variables are not measured (particularly with relation to the bypass valves) and this required to determine them based on modeling the heat and mass balance of the engine. Values for temperatures, mass flows and energy flows for different engine loads are provided in the \ref{sec:appendix:engines}, for both main and auxiliary engines.

\subsection{Electric power demand} \label{sec:met:electric}

The total electric power demand of the system is easily estimated as the sum of the power generated by the four auxiliary engines, that is directly measured on the electrical side of the generators and has a rather high accuracy.
\begin{equation}
P_{el,tot} = \sum_{1}^{4} P_{AE_i}
\end{equation}

The ship is equipped with a number of different systems, including lighting, navigational systems, pumps and compressors, etc. However, none of the individual contributions is directly measured, hence estimations of the energy demand of individual consumers need to be based on indirect measurements. In this paper only the contributions of the HVAC and the bow thrusters are identified separately from the remaining parts of the system. More details about the methods employed in this process are provided in Appendix REF.

\subsection{Heating and cooling systems} \label{sec:met:heat}

The case study ship is not provided with any means to directly measure the heat demand (e.g. steam flow meters), and it was hence necessary to provide estimations based on other measurements and assumptions. The heat demand is subdivided among the contributions of the HVAC preheater and reheater ($\dot{Q}_{HVAC,PH}$ and $\dot{Q}_{HVAC,RH}$), the hot water heater ($\dot{Q}_{HWH}$), the HFO heaters ($\dot{Q}_{HH}$) the tank heating systems for HFO and other liquids ($\dot{Q}_{HTH}$ and $\dot{Q}_{TH}$), the machinery space heating ($\dot{Q}_{MSH}$) and other tanks ($\dot{Q}_{OT}$). The heat generation is instead composed of the contribution of the exhaust gas boilers ($\dot{Q}_{EGB,ME,i}$ and $\dot{Q}_{EGB,AE,i}$), the heat recovery from the high temperature cooling systems ($\dot{Q}_{HTHR,j,i}$) and the boilers ($\dot{Q}_{AB}$).
 The heat demand and generation are summarized in equations \ref{eq:Qgen} and \ref{eq:Qdem}.
\begin{align}
\dot{Q}_{gen} & = \sum_{i=2,3} \dot{Q}_{EGB,ME,i} +  \sum_{i=1:4} \dot{Q}_{EGB,AE,i} + \sum_{j=ME,AE, i=1:4}\dot{Q}_{HTHR,j,i} + \dot{Q}_{AB} \label{eq:Qgen} \\
\dot{Q}_{dem} & = \dot{Q}_{HVAC,PH} + \dot{Q}_{HVAC,RH} + \dot{Q}_{HWH} + \dot{Q}_{HH} + \dot{Q}_{TH} + \dot{Q}_{G} + \dot{Q}_{OT} + 	\dot{Q}_{HTH} + \dot{Q}_{MSH} \label{eq:Qdem}
\end{align}

>>>>

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
%% The amsthm package provides extended theorem environments
\usepackage{amsthm}
%% Tthe amsmath package provides the "align" environment for nicer equation formatting
\usepackage{amsmath}
%% The package tabls helps in having nicer tables with higher spaces
\usepackage{tabls}
%% The package multirow allows merging together multiple cells vertifcally
\usepackage{multirow}
%% The package subcaption is for helping when using subfigures and their captions
\usepackage{subcaption}
%% This package is used for having nicer tables
\usepackage{booktabs}
% Package for improving table handling
\usepackage{array}
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
% Package for drawing pictures
\usepackage{tikz}
\usepackage{pgfplots}
% Package for the nomenclature
\usepackage{nomencl} 
\usepackage{ifthen} % Required for having groups in the nomenclature
\makenomenclature
%Redefines the command "nomgroup" for the nomeclature sot that several subgroups can be included in the nomenclature
\RequirePackage{ifthen}
\renewcommand{\nomgroup}[1]{%
	\ifthenelse{\equal{#1}{A}}{\item[\textbf{Acronyms}]}{%
		\ifthenelse{\equal{#1}{C}}{\item[\textbf{Constants}]}{%
			\ifthenelse{\equal{#1}{S}}{\item[\textbf{Subscripts}]}{%
				\ifthenelse{\equal{#1}{V}}{\item[\textbf{Variables}]}}}}}

@@

\author[EPFL]{Francesco Baldi}
\author[Linnaeus]{Fredrik Ahlgren}
\author[DTU,SP]{Tuong-Van Nguyen}
\author[LTU]{Marcus Thern}
\author[CTU]{Karin Andersson}

\address[EPFL]{Industrial Process Energy Systems Engineering (IPESE), \'{E}cole Polytechnique F\'{e}d\'{e}rale de Lausanne, 1950, Sion, Switzerland}
\address[Linnaeus]{Kalmar Maritime Academy, Linnaeus University, Kalmar, Sweden}
\address[SP]{Laboratory of Environmental and Thermal Engineering, Polytechnic School - University of S\~{a}o Paulo, S\~{a}o Paulo, Brazil}
\address[DTU]{Department of Mechanical Engineering, Technical University of Denmark, Lyngby, Denmark}
\address[LTU]{Energy Sciences, Lund University, Lund SE-22100, Sweden}
\address[CTU]{Department of Mechanics and Maritime Sciences, Chalmers University of technology, Gothenburg, Sweden}

\begin{abstract}
The shipping sector is today facing numerous challenges. Fuel prices are expected to increase in the medium-long term, and the recent decision of the IMO to require a reduction of shipping's greenhouse gas emissions by 50\% with respect to 2009 levels will important effort within this industry. In this context, passenger ships represent a small but increasing share of the industry. The complexity of the energy system of a ship where the energy required by propulsion is no longer the obvious main contributor to the whole energy use makes this kind of ship of particular interest for the analysis of how energy is converted from its original form to its final use on board.
To illustrate this, we analyzed the energy and exergy flow rates of a cruise ship sailing in the Baltic Sea based on a combination of available measurements from ship operations and of mechanistic knowledge of the system. The energy analysis allows identifying propulsion as the main energy user (46\% of the total) followed by heat (27\%) and electric power (27\%) generation; the exergy analysis allowed instead identifying the main inefficiencies of the system: While exergy is primarily destroyed in all processes involving combustion (76\% of the exergy destruction is generated by the combustion in the Diesel engines and in the oil-fired boilers), the other main causes of exergy destruction are the turbochargers (29\% of the remaining exergy destruction), the heat recovery steam generators (11\%), the steam heaters (11\%) the preheater in the accommodation heating systems (10\%), the sea water coolers (9\%), and the electric generators (6\%); the main exergy losses happen instead in the exhaust gas, mostly from the main engines (67\% of total losses), particularly from those not equipped with heat recovery devices. The improved understanding which derives from the results of the energy and exergy analysis can be used as a guidance to identify where improvements of the systems should be directed.
As a result of the energy flow analysis of the ship, in this paper we also propose a clustering of ship operations based on the concept of typical operational days. We found that, based on the operations of the case study, the use of five typical days, clustered based on the sum of mechanical and electrical power, and on heating demand, provides a good approximation of yearly ship operations and can hence be used for the design and optimization of the energy systems of the ship.

@@

\printnomenclature % Prints the nomenclature

@@

CO$_2$ emissions from shipping in 2012 amounted to a total of 949 million tonnes, contributing to 2.7\% of global anthropogenic CO$_2$ emissions \cite{Smith2014}. Although this contribution appears relatively low, the trend is that shipping will play an even greater role in the future due to the increased transport demand according to all IMO future scenarios \cite{Smith2014}. As an example, global transport demand has increased by 3.4\% in 2014, compared to a global gross product growth of 2.5\% the same year, which shows how shipping tends to rise even faster than global economy \cite{UNCTAD2015}. The OECD countries have reduced the CO$_2$ impact from shipping, but a larger amount has been moved to the non-OECD countries [3]. The fact that shipping needs to even further reduce its CO$_2$ emissions in the near future is essential for being able to achieve the goals of maintaining the climate below a 2-degree level in 2050 [4].

\nomenclature[A]{OECD}{Organisation for Economic Co-operation and Development}

In addition to considerations related to GHG emissions, an emission control area (ECA) is enforced in the Baltic Sea  by the International Maritime Organisation since January 2015. This ECA stipulates that the fuel used must not contain more than 0.1\% sulphur, therefore requiring the use of more expensive distillate fuels. More generally making shipping sustainable is a challenge that will demand growing attention by the shipping industry \cite{Andersson2016}.

\nomenclature[A]{GHG}{Greenhous gas}
\nomenclature[A]{ECA}{Emission Controlled Area}

@@

Altogether, these conditions present a challenge to the shipping companies that attempt to reduce their fuel consumption, environmental impact, and operative costs. A wide range of fuel saving solutions for shipping are available and partially implemented in the existing fleet, both from the design and operational perspectives; several specific studies have been conducted on these technologies, and a more detailed treatise would be out of the scope of this work. In this context, it has been acknowledged that the world fleet is heterogeneous, and measures need to be evaluated on a ship-to-ship basis \cite{Bouman2017}. In this process, a deeper understanding of energy use on-board of the specific ship is vital.

@@

Most authors focused on the propulsion part of the problem, as this is often the most relevant energy demand on-board. Shi et al. \cite{Shi2009} proposed a modeling approach for predicting ship fuel consumption of a cargo/passenger ferry; Theotokoats and Tzelepis \cite{Theotokatos2015} applied a similar procedure to the case of a Handymax product carrier, similarly to Tillig et al. \cite{Tillig2016}, who also added the dynamic element to their predictive model.

@@

% Data-driven analysis
Other authors filled the gap by both including electric energy demand, and by basing their work on measurements from actual ship operations. The work presented by Thomas et al. \cite{Thomas2010} and Basurko et al. \cite{Basurko2013} shows the application of energy auditing methods to fishing vessels. This approach represents a step forward towards improving the understanding of how energy is used on-board during actual ship operations.

@@

% Heat availability
Heat demand is rarely a subject of concern on-board ships, with some notable exceptions. This is due to a combination of generally low demand and high availability from the waste heat of the engines. In a previous publications by the authors \cite{Baldi2015a}, for instance, it is shown that in the case of a product tanker, although the heat demand was estimated to account for roughly 20\% of the total energy demand of the ship on a yearly basis, it only contributes to 4.1\% of the total fuel consumption (contribution of the auxiliary boilers), while the rest of the demand is fulfilled using waste heat.

%% WHR and heat availability
It should be noted that, however, much research effort has been devoted, especially in recent times, to the improvement of the efficiency of ship energy systems by recovering the waste heat available from the engines. With reference to different types of technologies, case studies, and designs,the several authors showed the existance of a quite significant potential for energy saving when WHR systems are employed, ranging from around 1\% for single-pressure steam cycles applied to two-stroke engines \cite{Theotokatos2012} to more complex systems based on ORCs (up to 10\%, \cite{Hountalas2012}) or including the cooling systems as a source of waste heat (over 10\%, \cite{Dimopoulos2012}).  The case of the installation of an ORC on-board of the vessel investigated in this study (see Section \ref{sec:met:case}) was presented in \cite{Mondejar2017}, showing a similar potential.

\nomenclature[A]{WHR}{Waste heat recovery}
\nomenclature[A]{ORC}{Organic Rankine cycle}

The potential uses for waste heat on-board are not limited to improving the efficiency of the power plant. Waste heat is commonly used for fulfilling on-board heat demand for spatial heating and freshwater generation \cite{Baldi2015a,Molland2011,Mccarthy1990};  Balaji et al. \cite{Balaji2012} proposed the use of waste heat for ballast-water systems ; Salmi et al. \cite{Salmi2017} suggested its use for adsorption refrigeration systems. A detailed review of potential uses for waste heat from marine engines is presented by Shu et al \cite{Shu2013}.

@@

% Exergy analysis
Exergy analysis provides a more accurate estimation of the potential for energy recovery on-board. The application of exergy analysis to the case of ship energy systems is, however, still limited. Dimopoulos et al. \cite{Dimopoulos2012} showed how the process optimizing the WHR system of a container ship can be more efficient if exergy efficiency, rather than energy efficiency, is used as the target of the optimization; Baldi et al. \cite{Baldi2015a} also analyzed the exergy flows on-board of a product tanker, showing that this allows having a more accurate understanding of what parts of the system show potential for improvement ; similar results were obtained by Marty et al. \cite{Marty2016}, who focused on the power plant of a cruise ship. Koroglu et al. \cite{Koroglu2017} made a step futher also including advanced exergy analysis in their study.

@@

In this paper, we present the application of energy and exergy analysis (described in Section \ref{sec:met:energyExergy}) to a cruise ship. This is done for one specific case study vessel, described in Section \ref{sec:met:case}; in Section \ref{sec:met:gathering} we present the specifics of the available information, in particular the measured data and the technical documentation; the specific assumptions and methods used for processing the available information into  energy and exergy flows are summarized in Section \ref{sec:met:processing}, while a more thorough description is provided in \ref{sec:app:method}.

@@

\begin{align}
	\sum_{\mathrm{in}} \dot{H}_{\mathrm{in}} &= \sum_{\mathrm{out}} \dot{H}_{\mathrm{out}}  \label{eqn:energyBalance}  \\
	\dot{H}_{\mathrm{fuel}} + \dot{H}_{\mathrm{air}} &= \sum_{{\mathrm{waste}}}\dot{H}_{{\mathrm{waste}}}+\dot{W}_{\mathrm{el}}+\dot{Q}_{\mathrm{heating}} \label{eqn:energyBalance2}
\end{align}

\nomenclature[V]{H}{Enthalpy}
\nomenclature[S]{in}{Inlet}
\nomenclature[S]{out}{Outlet}
\nomenclature[V]{Q}{Heat}
\nomenclature[S]{el}{Electric}

In Equation \ref{eqn:energyBalance2}, the left-hand side term represents, on a time rate basis, the energy associated with the fuel consumed in the boilers and engines ($\dot{H}_{\mathrm{fuel}}$) and the air used in the combustion processes ($\dot{H}_{\mathrm{air}}$). The right-hand side term denotes the power ($\dot{W}_{\mathrm{el}}$) and heat ($\dot{Q}_{\mathrm{heating}}$) required on-site (e.g. propulsion and fuel heating), and the heat discharged into the environment ($\sum_{{\mathrm{waste}}}\dot{H}_{{\mathrm{waste}}}$) with, for instance, the exhaust gases.

The energy flow associated with a material stream is calculated as the sum of the physical and chemical enthalpies, and kinetic and potential energies are neglected. The physical energy is taken as the relative enthalpy, as underlined in \cite{Kotas2013}, and the chemical energy is taken as the lower/higher heating value. The environmental conditions taken for the present analysis are the ambient pressure (1.01 bar) and seawater temperature (measured).

@@

\begin{align}
	\sum_{\mathrm{in}} \dot{E}_{\mathrm{in}} &= \sum_{\mathrm{out}} \dot{E}_{\mathrm{out}} +\dot{E}_{d} \label{eqn:exergyBalance} \\
	\dot{E}_{\mathrm{fuel}} + \dot{E}_{\mathrm{air}} &= \sum_{{\mathrm{waste}}}\dot{E}_{{\mathrm{waste}}}+\dot{E}_{W}+\dot{E}_{Q,\mathrm{heating}}+\dot{E}_{d}  \label{eqn:exergyBalance2}
\end{align}

\nomenclature[V]{E}{Exergy}
\nomenclature[S]{d}{Destroyed}
\nomenclature[S]{l}{Lost}
\nomenclature[S]{prod}{Product}

In Equation \ref{eqn:exergyBalance2} the left-hand side term represents, on a time rate basis, the exergy associated with the fuel and air. The right-hand side term denotes the exergy of the waste streams, the exergy transfers with heat and power, and the exergy destroyed in the ship system. Kinetic and potential exergies are neglected, and the exergy of a given material stream is derived as the sum of the physical and chemical exergies. The chemical exergy is calculated based on the reference environment of \cite{Szargut1989}, and its value is approximatively equal to the higher heating value for hydrocarbon fuels. The exergy destruction $(\dot{E}_d)$ can be calculated using the Gouy-Stodola theorem \cite{Kotas2013}.

The exergy balance may alternatively be formulated as:
\begin{equation}
\dot{E}_{\mathrm{p}} = \dot{E}_{\mathrm{fuel}} - \dot{E}_{\mathrm{d}} - \dot{E}_{\mathrm{l}} 
\end{equation}

where $\dot{E}_{\mathrm{p}}$ is called the exergy product, and corresponds to the desired output of the system, in exergy terms (for example, the power produced in an engine). $\dot{E}_{\mathrm{fuel}}$ denotes the exergy fuel, and represents the resources spent to drive the studied process (for instance, the fuel used in a combustion process). The last term $\dot{E}_{\mathrm{l}}$ corresponds to the losses of a system, such as the heat discharged into the environment with cooling water.

@@

\nomenclature[V]{$\varepsilon$}{Exergy efficiency}
\nomenclature[V]{$\lambda$}{Efficiency defect}
\nomenclature[V]{$\delta$}{Irreversibility share}
\nomenclature[S]{tot}{Total}

@@

The ship systems are summarized in Figure \ref{fig:shipSystems}. The propulsion system is composed of two propulsion lines, each made of two engines, a gearbox, and a propeller. The main engines are four W\"{a}rtsil\"{a} 4-stroke Diesel engines (ME) rated 5850 kW each. All engines are equipped with selective catalytic reactors (SCR) for NO$_X$ emissions abatement. Propulsion power is needed whenever the ship is sailing; however, it should be noted that the ship rarely sails at full speed, and most of the time it only needs one or two engines operated simultaneously.

@@

Auxiliary power is provided by four auxiliary engines (AE) rated 2760 kW each. Auxiliary power is needed on-board for a number of alternative functions, from pumps in the engine room to lights, restaurants, ventilation and entertainment for the passengers.

@@

The operational data was collected on-board from the ship's machine logging and surveillance system. The data sampling rate was selected to 15 minutes, automatically calculated as averages over the respective periods by the on-board data logging systems, as an acceptable compromise between accuracy and computational time required for the data processing. Data consistency was checked individually by creating a descriptive statistic and a histogram of each measurement. A filter was applied by setting up a pre-defined maximum and minimum value and eliminating from the analysis all data points where measurements were located outside these boundaries. This led to discarding XXX out of YYY data points.

@@

The main principles applied to the modeling are presented in the following sections and summarized in Figure \ref{fig:dataprocessing}, while a more thorough description and validation of the different approaches is presented in  \ref{sec:app:method}. The data are first processed by the main and auxiliary engines modules to derive the related energy flows. These are then used to calculate the electric power demand, the energy flows in the cooling systems and, consequently, the heat recovered in the HTHR systems. This information combined is used to determine the heating demand. Finally, all flows thus obtained are aggregated to determine energy and exergy flows and related performance indicators.

@@

The Diesel engines are the only source of electric and mechanical power in the system, subdivided in four main engines for propulsion and four auxiliary engines for on-board electric power generation. The energy and exergy analysis of the engines is not limited to the boundary of the system, but is brought to the detail of the different components that constitute the engine systems, as shown in Figure \ref{fig:DieselEngines}.

@@

\nomenclature[V]{$\omega$}{Rotational speed}
\nomenclature[V]{$\dot{m}$}{Mass flow}
\nomenclature[V]{$\eta$}{Energy efficiency}
\nomenclature[S]{des}{Design}
\nomenclature[A]{ME}{Main engine}
\nomenclature[A]{AE}{Auxiliary engine}
\nomenclature[V]{LHV}{Lower heating value}
\nomenclature[V]{frp}{Fuel rack position}

In Equations \ref{eqn:ME:eq1} to \ref{eqn:ME:eq3}, $\dot{m}_{fuel}$, $frp$, $\omega$, $\eta_{ME}$, and $LHV$ represent the fuel flow to the engine, the fuel rack position, the engine speed, the engine efficiency and the lower heating value of the fuel, respectively, and the $des$ subscript refers to design conditions.

@@

For both auxiliary and main engines, part of the relevant internal variables are not measured (particularly with relation to the bypass valves) and this required to determine them based on modeling the heat and mass balance of the engine. Values for temperatures, mass flows and energy flows for different engine loads are provided in the \ref{sec:app:met:engines}, for both main and auxiliary engines.

\subsubsection{Electric power demand} \label{sec:met:electric}

The total electric power demand of the system ($P_{el,tot}$) is easily estimated as the sum of the power generated by the four auxiliary engines ($P_{AE}$), that is directly measured on the electrical side of the generators and has a rather high accuracy.
\begin{equation}
P_{el,tot} = \sum_{1}^{4} P_{AE_i}
\end{equation}

\nomenclature[V]{P}{Power}

The ship is equipped with a number of different systems, including lighting, navigational systems, pumps and compressors, etc. However, none of the individual contributions is directly measured, hence estimations of the energy demand of individual consumers need to be based on indirect measurements. In this paper only the contributions of the HVAC and the bow thrusters are identified separately from the remaining parts of the system. More details about how these contributions were separated are provided in \ref{sec:app:met:electric}.

\subsubsection{Heating and cooling systems} \label{sec:met:heat}

The main view of the ship heating and cooling systems is presented in Figure \ref{fig:ship:general}. The case study ship is not provided with any means to directly measure the heat demand (e.g. steam flow meters) and cooling flows, and it was hence necessary to provide estimations based on other measurements and assumptions.

The ship's cooling systems are designed similarly to most ships and divided in two cooling systems operated at two separate temperature levels: the high temperature (HT) cooling systems (70-90$^o$C) and the low temperature (LT) cooling systems (40-60$^o$C). The cooling systems provide the necessary cooling to the engines, to the steam systems and to all other components on-board.

\nomenclature[A]{HT}{High temperature}
\nomenclature[A]{LT}{Low temperature}

All values not directly measured in the cooling systems were calculated based on mass and energy balances. Mass flows were determined based on measurements of the pressure in the system and on the performance curves of the cooling pumps as provided in the engine project guides.

The heat demand is subdivided among the contributions of the HVAC preheater ($\dot{Q}_{HVAC,PH}$) and reheater ($\dot{Q}_{HVAC,RH}$), the hot water heater ($\dot{Q}_{HWH}$), the HFO heaters ($\dot{Q}_{HH}$) the tank heating systems for HFO ($\dot{Q}_{HTH}$) and other liquids ($\dot{Q}_{TH}$), the machinery space heating ($\dot{Q}_{MSH}$) and other tanks ($\dot{Q}_{OT}$).
